---
title: 'Statistical report: Discontinuation of guideline-recommended HF treatments after a HF hospitalization'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
---

\newpage 
\tableofcontents 
\listoftables
\listoffigures
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

SHFDB3, https://kiheartfailure.github.io/shfdb3/, v 3.2.4. 

## Inclusion/exclusion criteria

```{r flow}
default_kable(flow, caption = "Flowchart")
```

First patient in: `r min(rsdata$shf_indexdtm)` and last patient in: `r max(rsdata$shf_indexdtm)`.  

The median age (IQR) is `r rsdata %>% summarise(med = fn(median(shf_age), dig = 1),
                                             q1 = fn(quantile(shf_age, probs = 0.25), dig = 1),
                                             q3 = fn(quantile(shf_age, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r rsdata %>% count(shf_sex) %>%
  mutate(perc = fn(n / sum(n) * 100, 1)) %>%
  filter(shf_sex == "Female") %>%
  pull(perc)`% females.    

## Created variables 

```{r ddr}
default_kable(metalm, caption = "Medications from DDR", scale_down = F)
```

Prior medication is defined as 1 dispensed prescription within 6 months to 1 day prior to index date. 
A sensitivity analysis is performed defining prior medication as 2 dispensed prescriptions, with at least 1 day between, 
within 8 months to 1 day prior to index date. 

Post medication is defined as 1 dispensed prescription from index date until 6 months after index date. 

```{r npr}
default_kable(metaout, caption = "Additional outcomes from NPR")
```

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

## Baseline characteristics

```{r, child = "../src/tab1.Rmd"}

```

## Missing data

Missing data was imputed with multiple imputation (n = 10) using mice [@mice]. 
Variables included in the model are indicated in 
Table \ref{tab:tab1mra}. 

## Treatment discontinuation/initiation 

A logistic regression is performed separately for the 3 EF groups for patients 
with the respective medications prior to the hospitalization/visit, with discontinuation 
(for MRA, RASi, Bbl)/initiation (Loopdiuretics) as the outcome. 
Adjustment is performed for variables indicated in \ref{tab:tab1all}.  

```{r, child = "../src/treatdisclogreg.Rmd"}

```

\clearpage 

## Repeated hospitalizations

The barplot presents the number HF hospitalizations, defined as main diagnosis in NPR, 
including the index hospitalization, 
from index until the respective time point vs discontinuation of the respective treatment. 
The treatment is defined as 1 dispensed prescription within 6 months after 
the respective time point. Patients without time point + 6 months follow-up 
are excluded (as well as patients without prior treatment and patients that are 
not hospitalized at index (controls)). E.g., when looking 
at 1 year it is number of HF hospitalizations within 1 year from index vs 
treatment within 1-1.5 years after index and excluding patients without 1.5 years 
follow-up and without the treatment prior to index. 

```{r, child = "../src/rephosp.Rmd"}

```

\clearpage 

## Predictors of discontinuation/initiation 

A logistic regression is performed separately for the 3 EF groups for patients 
with the respective medications prior to the hospitalization, with discontinuation/initiation as the outcome. 
Only patients hospitalized for HF are included (so the controls are dropped from this analysis). 

```{r, child = "../src/predtreatdisc.Rmd"}

```

### Assumptions

Outliers were investigated with Cook's distance and multicollinearity 
with the variance inflation factor.
=> No action deemed necessary. 

## Outcomes after discontinuation

Data were censored at 2019-12-31, death (where applicable), emigration or 3 years. 

Cox proportional hazards regressions were 
used to model the time to first event analysis and negative 
binomial regression including the log of time as an offset in the model for 
the repeated events. The models were partly crude and partly adjusted for variables 
indicated in Table \ref{tab:tab1all} and are performed separately in the respective EF 
groups and on those with the respective prior treatments. 

The median (min-max) follow-up among the cases is `r rsdata %>% 
filter(sos_location == "HF in-patient") %>%
summarise(med = fn(median(sos_outtime_death3y / 365), dig = 1),
                                             min = fn(min(sos_outtime_death3y / 365), dig = 1),
                                             max = fn(max(sos_outtime_death3y / 365), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", min, "-", max, ")")) %>%
                                   pull(out)` years for a total of 
                                   `r rsdata %>% 
filter(sos_location == "HF in-patient") %>% summarise(sumpy = fn(sum(sos_outtime_death3y) / 365, dig = 0)) %>%
                                   pull(sumpy)` patient-years of follow-up.


```{r, child = "../src/treatdisccoxreg.Rmd"}

```

### Assumptions

The proportional hazards assumption was investigated using the scaled Schoenfeld 
residuals and linearity for the continuous variable age for the outcome CV mortality/first HF hospitalization within 3 years. 
=> No action deemed necessary. 

\clearpage
\newpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/treatdiscontinuation. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
